\documentclass[a4paper] {article}

\addtolength{\hoffset}{-1.27cm} 
\addtolength{\textwidth}{2.54cm}

\pagestyle{headings}

\usepackage{amsmath}
\usepackage{verbatim}
\usepackage{longtable}
\usepackage{calc}
\usepackage{graphicx}
\usepackage{fancyvrb}

% always load hyperref last since it redefines many LaTeX commands
% also you need to have hyperref turned off if you are using YAP to view .dvi files
\usepackage[colorlinks=true,bookmarks=true]{hyperref}
\setlongtables

%\setlength{\topmargin}{-20mm}
%\setlength{\oddsidemargin}{-2.4mm}
%\setlength{\evensidemargin}{-2.4mm}
%\setlength{\textwidth}{164mm}
%\setlength{\textheight}{256mm}
%\setlength{\columnsep}{8mm}
%\setlength{\columnseprule}{0.4pt}

\parindent0mm
\parskip2ex plus1ex minus1ex



\author{Peter Taylor}
\date{8 November 2014}
\sloppy

% define various commands for the standard constructs we will use in this document

\title{SPI User Guide (DRAFT)}
\begin{document}
\maketitle
\begin{abstract}
This document is a user guide for the Sartorial Programming Interface (SPI).
We describe the motivation for the product and demonstrate how to use it.
\end{abstract}
\tableofcontents
\section{Introduction}
This document is a user guide for the Sartorial Programming Interface (SPI).
We will describe the motivation for the product and demonstrate how to use it.

The purpose of SPI is to allow you to specify an interface which allows you to wrap an existing set of C++ libraries (or C libraries since C is a sub-set of C++) in a clean and well-documented interface.
The wrapping code is generated by a number of separate code generation tools:
\begin{itemize}
\item spcl: Generates C++ code which supports serialization and logging.
\item spxl: Generates an Excel interface.
\item sppy: Generates a Python interface.
\item spcli: Generates a .NET interface via C++/CLI language.
\item sptex: Generates a tex document.
\end{itemize}
We intend to add more code generators for different target platforms over the course of time.

SPI has a configuration language which is used to specify the interface.
This language is independent of the various target platforms (e.g. Excel, Python) and the language uses many C++ keywords.
As a result the configuration language looks a bit like a mixture of C++ header and source files.

The configuration files are designed to exist independently of the underlying C++ library that you are wrapping.
Our goal is to be able to wrap any C++ library in a non-intrusive fashion.
In other words the C++ library exists, and we define an interface to it so that the C++ library is made available to platforms that would be used to develop applications.
We do not use the approach (for example) of editing the C++ libraries own header files with special annotations, and then interpreting those annotations.

The document is broken down into the following further sections which will expand in ever increasing detail on the SPI configuration language:
\begin{itemize}
  \item Executive summary. Describes standard use cases, feature set of SPI, possible future developments, plus examples of use.
  \item High level user guide. 
  A high level user guide for using the code generation tools and also introduces the main components of SPI.
%  \item Detailed user guide. 
%  More detailed description of all the language features in SPI configuration files.
  \item SPI reference. This is the full reference for the use of SPI, together with examples of each SPI syntax element.
\end{itemize}

\newpage
\input{spi-user-guide/executive-summary.tex}

\newpage
\input{spi-user-guide/high-level.tex}

\newpage
\input{../../sample/spi-examples.tex}

\newpage
\input{spi-user-guide/reference.tex}

%\newpage
%\begin{thebibliography}{9}
%\bibitem{boost} http://www.boost.org \textbf{Boost C++ Libraries.}
%\bibitem{swig} http://www.swig.org. \textbf{Simplified Wrapper and Interface Generator.}
%\bibitem{functor} http://en.wikipedia.org/wiki/Function\_object.
%\textbf{Wikipedia entry for functor.}
%\end{thebibliography}

\end{document}
